\documentclass[11pt]{ucthesis}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algcompatible}
\usepackage{algpseudocode}
\usepackage{siunitx}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{subcaption}
\usepackage{float}
\sisetup{
  group-separator = {,}, 
  group-minimum-digits = 4,
  table-number-alignment = right, 
  group-digits          = integer
}


\def\dsp{\def\baselinestretch{2.0}\large\normalsize}
\dsp

\makeatletter
\renewcommand{\endabstract}{\par\end{alwayssingle}}
\makeatother

\begin{document}
\title{Effective Carbon Intensity Forecasting Methods for Grid and Carbon-Aware Optimizations}
\author{Mathew Nellariyil Joseph}
\degreeyear{2026}
\degreemonth{June}
\degree{MASTER OF SCIENCE}
% \chair{Professor Samuel}
\numberofmembers{4} %% (including chair) possible: 3, 4, 5, 6
\committeememberone{Professor Abel Souza, Chair}
\committeemembertwo{Professor Andi Quinn}
\committeememberthree{Professor Peter Alvaro}
\deanlineone{Peter Biehl}
\deanlinetwo{Vice Provost and Dean of Graduate Studies}
\deanlinethree{}
\field{Computer Science and Engineering}
\campus{Santa Cruz}

\begin{frontmatter}
\maketitle
\copyrightpage
\tableofcontents
\listoffigures
\listoftables

\begin{abstract}
\input{abstract}
\end{abstract}

\begin{acknowledgements}
  % TODO, probably thank Nautilus as well
  I would like to thank everyone who have been instrumental in my life and to this thesis.
  
  First and foremost, I would like to thank my advisor Dr. Abel Souza.
  It has long been my aspiration to have my computer science background help preserve mother nature, and he introduced me to a meaningful path for doing so.        
  His mentorship has been invaluable from the beginning of my Master's program; he taught me how to conduct impactful research and inspired in me a lasting commitment to the pursuit of carbon optimality.  
  
  I deeply grateful to my committee members, Dr. Andi Quinn and Dr. Peter Alvaro, for their feedback and guidance on my thesis. 
  Beyond the thesis, their expertise and instruction during my undergraduate studies were formative in helping me discover my research interests.
  
  I extend thanks to my lab mates Son Nguyen, Eugene Chou, and Omkar Lonkar who generously shared their experiences and insights, shaping me into a more capable researcher. 
  
  I want to thank my family for their steadfast love and support throughout my academic journey. 
  During my most demanding periods, they were my backbone and source of strength that carried me through every challenge. 
  
  I would also like to thank my close friends Arshad Mohammad, Alexander Colloredo-Mansfeld, Anoushka Singal, and Jack Lund. 
  Their enduring friendship and encouragement have meant more than I can express.
  
  Finally, I would like to thank Scouts of America, Troop 199/155, and Mr. Henry Allen for showing me the beauty in the great outdoors. I want to preserve these natural wonders for others to have the same experience. 
\end{acknowledgements}
\end{frontmatter}

\chapter{Introduction}
\label{chap:introduction}
\input{introduction}

\chapter{Background}
\label{chap:background}
\input{background}

\chapter{Methodology}
\label{chap:methodology}
\input{method}

\chapter{Evaluation}
\label{chap:evaluation}
\input{evaluation}

\chapter{Related Work}
\label{chap:related}
\input{related}

\chapter{Conclusion}
\label{chap:conclusion}
\input{conclusion}

\nocite{*}
\bibliographystyle{plain}
\bibliography{report}

\end{document}